---
title: "Bioinformatics Class 11"
author: "Pallas Duong"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Section 1
##PDB statistics 

Download CSV file from PDB database from <https://www.rcsb.org/stats/summary>. Read this into R and determine fraction of X-ray structure 

```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Lets calculate something 
```{r}
percent <- (pdb.stats$Total/sum(pdb.stats$Total))*100
names(percent) <- pdb.stats$Experimental.Method
percent 
```

#Section 3
##3.1: Getting started 
```{r}
library("bio3d")
```

##3.2: Bio3D functions and their typical usage
```{r}
example(plot.bio3d)
```

##3.3: Working with individual PDB files
```{r}
#Reading PDB file data into R
pdb <- read.pdb("1hsg")
pdb
```

```{r}
#Find the attributes
attributes(pdb)

head(pdb$atom)

#Print the subset of $atom data for the first 2 atoms
pdb$atom[1:2, c("eleno", "elety", "x", "y", "z")]

```

```{r}
#Individual $atom records
pdb$atom$elety[1:2]

plot.bio3d(pdb$atom$b[pdb$calpha], sse = pdb, ylab = "B-factor")
```

```{r}
#Print a summary of the coordinate data in $xyz
pdb$xyz
```

```{r}
#Exame the row and column dimensions
dim(pdb$xyz)
```

```{r}
#Print coordinates for the first 2 atoms
pdb$xyz[1, atom2xyz(1:2)]
```

3.5: Atom selection
```{r}
#Select all C-alpha atoms 
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

```{r}
#Print details of the first few selected atoms 
head(pdb$atom[ca.inds$atom,])
```

```{r}
#Select xyz coordinates 
head(pdb$xyz[, ca.inds$xyz])
```

```{r}
#Select chain A
a.inds <- atom.select(pdb, chain = "A")

#Select C-alphas of chain A
ca.inds2 <- atom.select(pdb, "calpha", chain = "A")

#we can combine multiple selection criteria to return their intersection 
cab.inds <- atom.select(pdb, elety = c("CA", "CB"), chain = "A", resno = 10:20)
```

Write pdb
```{r}
inds <- atom.select(pdb, "calpha")
inds
head(pdb$atom[inds$atom,])
```
```{r}
#Trim
ca.pdb <- trim.pdb(pdb,inds = inds)
ca.pdb

```
 
```{r}
#Write pdb
write.pdb(ca.pdb, file = "calpha_only.pdb")
```

#Section 4
##4.2: Aligning mulitple structures
```{r}
#Download some example PDB file
ids <- c("1TND_B", "1AGR_A", "1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <-  get.pdb(ids, split = TRUE)
```

```{r}
#Etract and align the chains we are interested in 
pdbs <- pdbaln(files, fit = TRUE)

#Print to screen a summary of the 'pdbs' object
pdbs
```

```{r}
#Access the first 5 rows, and 8 columns 
pdbs$ali[1:5, 1:8]
```

```{r}
#Associated residues numbers
pdbs$resno[1:5, 1:8]

```

```{r}
#Inspect alignmnet for gaps
gaps <- gap.inspect(pdbs$ali)

#Indices of non-gap columns 
head(gaps$f.inds)

#Access non-gap columns 
pdbs$ali[,gaps$f.inds]
```

##4.3: Basic structure analysis
```{r}
#Calculate sequence identity
seqidentity(pdbs)

#Calculate RMSD
rmsd(pdbs)
rd <- rmsd(pdbs)

#Clustering 
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k= 3)

#Plot results as dendrogram
hclustplot(hc, k = 3)
```

##4.4: principal component analysis 
```{r}
#Perform PCA
pc <- pca(pdbs)
plot(pc, col = grps)
```

#Section 5
##5.1: Search and retrieve Adenylate kinase structures 

```{r, cache= TRUE}
aa <- get.seq("1ake_A")

#blast or hmmer search
b <- blast.pdb(aa)
```

```{r}
#Plot a summary of search results 
hits <- plot(b)
```

```{r}
head(hits$pdb.id)
```

```{r}
#Fetch PDBs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)

#Align structures
pdbs <- pdbaln(files)

#Vector containing PDB
ids <- basename.pdb(pdbs$id)

#Draw schematic alignmnet 
plot(pdbs, labels = ids)
```

##5.2: Sequence conservation analysis 
```{r}
#Calculate sequence consevation
cons <- conserv(pdbs, method = "entropy22")

#SSE annotations 
sse <- pdbs2sse(pdbs, ind = 1, rm.gaps = FALSE)

#Plot conservation per residue 
plotb3(cons, sse=sse, ylab = "Sequence entropy")
```

##5.4: Principal component analysis 
```{r}
#Find invariant core 
core <- core.find(pdbs)

#superimpose all structures to core 
pdbs$xyz <- pdbfit(pdbs, core)

#Perform PCA
pc.xray <- pca(pdbs)
```

```{r}
#Calculate RMSD
rd <-  rmsd(pdbs)

#Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k = 3)

plot(pc.xray, 1:2, col = "grey50", bg = grps.rd, pch = 21, cex =1)

```

